---
title: "BEA data"
author: "Vlad"
date: "2022-10-15"
output: html_document
---


### loading the packages

```{r setup, include=FALSE}
# load our packages for today
library(tidyverse)
library(readxl)
library(janitor)
library(curl)
library(lubridate)
library(usethis)
knitr::opts_chunk$set(echo = TRUE)

```

### importing the data

```{r}
url <- "https://apps.bea.gov/national/Release/XLS/Survey/Section1All_xls.xlsx"
```

```{r}
destfile <- "Section1All_xls.xlsx"
curl::curl_download(url, destfile)
```

### choosing the sheet
```{r}
read_excel(destfile,
           sheet = "T10103-A")
```

### wrangling the data
# skip the first 7 rows
```{r}
read_excel(destfile,
           sheet = "T10103-A",
           skip = 7) %>%
  head()
```

# remove a column called Line
```{r}
read_excel(destfile, sheet = "T10103-A", skip = 7) %>%
  select(-Line) %>%
  # view the first 10
  head(10)
```

```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = `...2`, code = 2) %>%
  head()
```
# change everything that contains A to numeric
```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(contains("A"), as.numeric))
```
# making long data
```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
  ) %>%
  pivot_longer(-account:-code, names_to = "date", values_to = "quarter_percentchange")
```
# format the dates

```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
  ) %>%
  pivot_longer(-account:-code, names_to = "date", values_to = "quarter_percentchange") %>%
  mutate(date = parse_date_time(date, orders = "Yq")) %>%
  tail()
```
```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code =2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
  ) %>%
  pivot_longer(-account:-code, names_to = "date", values_to = "quarter_percentchange") %>%
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date))
```
# filtering 
```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
  )%>%
  pivot_longer(-account:-code, names_to = "date", values_to = "quarter_percentchange") %>%
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>%
  filter(account == "Durable goods")
```
# filtering the date and quarterly percent change
```{r}

read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
  ) %>%
  pivot_longer(-account:-code,
               names_to = "date",
               values_to = "quarter_percentchange") %>%
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>%
  filter(quarter_percentchange < 1, 
         date >= "2020-01-01",
         account == "Durable goods")
```
```{r}
read_excel(destfile, sheet = "T10103-A",
           skip = 7) %>%
  select(-1) %>%
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
    ) %>%
  pivot_longer(-account:-code,
               names_to = "date",
               values_to = "quarter_percentchange") %>%
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>%
  group_by(account) %>%
  slice(1, n()) %>%
  head()
```
  
# arrange in descending order
```{r}
read_excel(destfile, sheet = "T10103-A", 
    skip = 7) %>% 
  select(-1)  %>% 
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
    ) %>% 
  pivot_longer(-account:-code, 
               names_to = "date", 
               values_to = "quarter_percentchange") %>% 
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>% 
  filter(account == "Durable goods")  %>% 
  arrange(desc(date)) %>% 
  head()
```

# rearrange columns
```{r}
read_excel(destfile, sheet = "T10103-A", 
    skip = 7) %>% 
  select(-1)  %>% 
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
    ) %>% 
  pivot_longer(-account:-code, 
               names_to = "date", 
               values_to = "quarter_percentchange") %>% 
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>%  
  filter(account == "Durable goods")  %>% 
  arrange(desc(date)) %>% 
  relocate(date, everything()) %>% 
  head()

```
# naming the final product

```{r}
durable_goods_quarter_percent_change <- read_excel(destfile, sheet = "T10103-A", 
    skip = 7) %>% 
  select(-1)  %>% 
  rename(account = 1, code = 2) %>%
  mutate(
    across(c(-account, -code), as.numeric)
    ) %>% 
  pivot_longer(-account:-code, 
               names_to = "date", 
               values_to = "quarter_percentchange") %>% 
  mutate(date = parse_date_time(date, orders = "Yq"),
         date = ymd(date)) %>%  
  filter(account == "Durable goods")  %>% 
  arrange(desc(date)) %>% 
  relocate(date, everything()) 

durable_goods_quarter_percent_change %>% 
  distinct(account, .keep_all = T)
```

### mutate function
<https://dplyr.tidyverse.org/reference/mutate.html> ## link for the mutate function
### mutate function allows us to add new variables while preserving the existing varibles in the dataset we are working with