---
title: "HIP"
author: "Vlad"
date: "2022-10-15"
output: html_document
---

### 1. Set up Chunk

```{r setup, include=F}

library(tidyverse)
library(readxl)
library(janitor)
library(scales)
library(lubridate)
library(timetk)
library(riingo)
library(tidyquant)
library(gt)
library(gtExtras)
library(reactable)
library(reactablefmtr)
library(geomtextpath)
library(curl)
library(readxl)
library(usethis)
library(tidytext)
library(pdftools)
library(corrr)

knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, comment = NA)

```

### 2. Import the data

```{r}
data_fmhpi <- read.csv("https://www.freddiemac.com/fmac-resources/research/docs/fmhpi_master_file.csv")

head(data_fmhpi)
summary(data_fmhpi)
```


### 3. Wrangle
create a new variable Date in y-m-d format
```{r}
data_fmhpi$date <- as.Date(with(data_fmhpi, paste(Year, Month, 01, sep="-")), "%Y-%m-%d")

head(data_fmhpi)
```
filter to observations > 1999-12-01
```{r}

data_fmhpi <- data_fmhpi %>% filter(date > "1999-12-01")

```
only keep the data for AK, GA, CA, CO, TX

```{r}
data_fmhpi <- data_fmhpi %>% filter(GEO_Name %in% c("AK", "GA", "CO", "CA", "TX"))
head(data_fmhpi)
```

changing the name of the column GEO_Name to State
```{r}

colnames(data_fmhpi)[4] = "State"
head(data_fmhpi)

```
```{r}
data_fmhpi <- data_fmhpi %>% select(4, 6, 8)
head(data_fmhpi)
```
create a new column "percent change" 
save the new object as a new dataframe "data_fmhpi1
```{r}
data_fmhpi1 <- data_fmhpi %>%
  arrange(date) %>%
  mutate(percent_change = (Index_NSA/lag(Index_NSA)-1) * 100)
head(data_fmhpi1)
```
### 4. Chart the trend over time in HPI for 5 states

```{r}
data_fmhpi1 %>%
  ggplot(aes(x = date, y = Index_NSA, fill = State)) +
  geom_col(
    show.legend = F
  ) +
  labs(title = "HPI trend over time",
       y = "",
       x = "",
       fill = "") +
  #scale_y_continuous(labels = percent_format(scale = 1),
  #                   breaks = pretty_breaks(n = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5) ) +
  facet_wrap(~State,
             ncol = 2)

```
HPI for AK has a steady slow increase from 2000-2022
HPI for CO has a steady increase from 2000 to 2012 and then we see an increase at a higher rate with a peak around 2022
HPI for TX has a relatively steady and slow increase from 2000 to 2012 and then we see an increase at a higher rate from 2012 to 2022
HPI for CA has a fast growing increase with a peak around 2007 and then we have a decreases from 2007 to about 2012 with a steady increase and a peak around 2021
HPI for GA has a slow increase until 2008 and a slow increase from 2008 to 2012, and a slow increase from 2012 to 2022

```{r}
data_fmhpi1 %>% 
  ggplot(aes(x = date, y = Index_NSA, color = State)) +
  geom_line() +
  labs(title = "HPI trend over time", 
       y = "", 
       x = "",
       fill = "") +
 # scale_y_continuous(labels = percent_format(scale = 1),
  #                   breaks = pretty_breaks(n = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5) ) +
  facet_wrap(~State,
             ncol = 2)
```


chart options:

```{r}
data_fmhpi1 %>%
  ggplot(aes(x = date, y = percent_change, fill = State)) +
  geom_line(show.legend = F,
             color = "steelblue") +
  labs(title = "Percent Change HPI from Previous Quarter", 
       y = "", 
       x = "",
       fill = "") +
  scale_y_continuous(labels = percent,
                     breaks = pretty_breaks(n = 5)) +
 
  theme_minimal() +
  theme(plot.title = element_text(hjust = .7)) +
  facet_wrap(~State,
             ncol = 2)
```
```{r}
data_fmhpi1 %>%
  ggplot(aes(x = date, y = percent_change, fill = State)) +
  geom_col(fill = "steelblue",
           width = 15) +
  labs(title = "Percent Change HPI from Previous Quarter", 
       y = "", 
       x = "",
       fill = "") +
  scale_y_continuous(labels = percent,
                     breaks = pretty_breaks(n = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .75)) + 
  facet_wrap(~State,
             ncol = 2)
```
### 5. Chart the trend over time in annual percent change in HPI for 5 states
```{r}
data_fmhpi1 %>% 
  ggplot(aes(x = date, y = percent_change, color = State)) +
  geom_line() +
  labs(title = "Annual Percent Change in HPI over time", 
       y = "", 
       x = "",
       fill = "") +
  scale_y_continuous(labels = percent_format(scale = .5),
                     breaks = pretty_breaks(n = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .7) ) +
  facet_wrap(~State,
             ncol = 2)
```

```{r}
data_fmhpi1 %>% 
  ggplot(aes(x = date, y = percent_change, fill = State)) +
  geom_col() +
  labs(title = "Annual Percent Change in HPI over time", 
       y = "", 
       x = "",
       fill = "") +
  scale_y_continuous(labels = percent_format(scale = .5),
                     breaks = pretty_breaks(n = 5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .7) ) +
  facet_wrap(~State,
             ncol = 2)
```
Annual percent change in AK is positive from the year of 2000 to 2016 and then we have a switch a slow decrease to a negative annual percent change from 2016 to 2022
Annual percent change in CO starts at a low positive percentage and then decreases at a high rate until 2006 and then we see soft increases and decreases from 2006 to 2022
Annual percent change in TX is close to 0 from 2000 to 2008 and then we see an increase at a high rate from 2008 to 2022 with a peak at 2011
Annual percent change in CA starts of with low negative numbers from 2000-2001 with a fast increase from 2001 to 2006 and then a fast decrease from positive to negative change from 2006 to 2012 and then we see a steady increase from negative change to positive change from 2012 to 2022
Annual percent change in GA starts of with change close to 0 from 2000 to 2008 and then a steady decrease with small deviations from 2008 to 2022
### 6. Distribution of percent change in HPI for 5 states
```{r}
data_fmhpi1 %>% 
  ggplot(aes(x =  percent_change, fill = State)) +
  geom_histogram(
    show.legend = F
  ) +
  labs(title = "Histrograms of Annual Percent Changes in HPI", 
       y = "", 
       x = "",
       fill = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .1),
        axis.text.x = element_blank()) +
  facet_wrap(~State)
```
Distribution of annual percent changes in HPI in AK is left skewed with a lot of deviations
Distribution of annual percent changes in HPI in CA is bimodal with some deviations
Distribution of annual percent changes in HPI in CO is left skewed with deviations
Distribution of annual percent changes in HPI in GA in bimodal with deviations
Distribution of annual percent changes in HPI in TX is bimodal with deivations

### 7. Chart the relationship
first we wrangle data
then plot scatterplot

```{r}
data_fmhpi1 %>% 
  filter(State %in% c("GA", "TX")) %>%
  select(-Index_NSA) %>%
  pivot_wider(names_from = State,
              values_from = percent_change)
```
```{r}
data_fmhpi1 %>% 
  filter(State %in% c("GA", "TX")) %>%
  select(-Index_NSA) %>%
  pivot_wider(names_from = State,
              values_from = percent_change) %>% 
  ggplot(aes(x =  GA, y = TX, 
             color = year(date) %>% as_factor())) +
  geom_point(
    #color = "purple",
    show.legend = T
  ) +
  geom_smooth(
    method = "lm",
    se = F,
    color = "pink",
    size = .5
    
  ) +
  labs(title = "GA v. TX", 
       y = "TX", 
       x = "GA")
  # scale_y_continuous(labels = percent_format(scale = 1),
  #                    breaks = pretty_breaks(n = 10)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = .5),
        axis.text.x = element_blank())
```
### 8. create a table displaying the rolling 3-month average HPI for the last 6 months

```{r}
data_fmhpi1 %>%
  summarise_by_time(.date_var = date,
                    .by = "3 months",
                    avg_HPI = mean(Index_NSA)) %>%
  tail()


```

### 9 import data from FRED on CPI inflation since 1999 and join to existing data frame
```{r}
cpi_core <- 
  "CPIAUCSL" %>% 
  tq_get(get = "economic.data", from = "1999-01-01") %>%
  tail()

cpi_core %>%
  summarise_by_time(.date_var = date,
                    .by = "3 months",
                    avg_price = mean(price)) %>%
  tail()
cpi_core1 <- cpi_core %>%
  arrange(date) %>%
  mutate(percent_change_cpi = (price/lag(price)-1) * 100)
head(cpi_core1)

joint_data <-
  cpi_core1 %>%
  select(-symbol) %>%
  left_join(
    data_fmhpi1,
    by = "date"
  )
  
```

### 10 which states appear to be highly correlated with overall inflation data since 1999?
rolling correlation vs overall correlation
chart to support answer

```{r}
joint_data %>% 
correlate( quiet = TRUE) %>%
gt()

```
```{r}
joint_data %>%
  filter(State == "AK") %>%
  correlate(quiet = TRUE) %>%
  gt()
```
Correlation between annual change in CPI and annual change in HPI for AK is -0.80377512 which is a high negative correlation.
```{r}

joint_data %>%
  filter(State == "CA") %>%
  correlate(quiet = TRUE) %>%
gt()
```
Correlation between annual change in CPI and annual change in HPI for CA is 0.49137226 which is relatively weak positive correlation
```{r}

joint_data %>%
  filter(State == "CO") %>%
  correlate(quiet = TRUE) %>%
gt()
```
Correlation between annual change in CPI and annual change in HPI for CO is -0.32280093 which is a weak negative correlation
```{r}
joint_data %>%
  filter(State == "GA") %>%
  correlate(quiet = TRUE)%>% 
gt()
```
Correlation between annual change in CPI and annual change in HPI for GA is -0.47206470 which is a relatively weak negative correlation
```{r}
joint_data %>%
  filter(State == "TX") %>%
  correlate(quiet = TRUE) %>%
gt()

```
Correlation between annual change in CPI and annual change in HPI for TX is 0.54156440 which is aa relatively weak positive correlation
